Evaluating machine translation for assimilation via a gap-filling task

### 2015-04-03 2015-eamt-assim/2015-eamt-assim.tex
Deanonimized
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59615 72bbbca6-d526-0410-a7d9-f06f51895060

\author{\blankout{Ekaterina Ageeva}\\
  \blankout{School of Linguistics}\\
  \blankout{Higher School of Economics}\\
  \blankout{Moscow, Russia}\\
  \blankout{{\tt evageeva\_2@edu.hse.ru}}\\[2ex]
  \blankout{\textbf{Francis M. Tyers}}\\
  \blankout{HSL-fakultetet}\\ 
  \blankout{UiT Norgga \'{a}rktala\v{s} universitehta} \\
  \blankout{9017 Romsa, Norway} \\
  \blankout{{\tt francis.tyers@uit.no}}
  \And
  \blankout{Mikel L. Forcada}\\
  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
  \blankout{Universitat d'Alacant, Spain} \\
  \blankout{{\tt mlf@dlsi.ua.es}}  \\[2ex]
  \blankout{\textbf{Juan Antonio P\'{e}rez-Ortiz}} \\
  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
  \blankout{Universitat d'Alacant, Spain} \\
  \blankout{{\tt japerez@dlsi.ua.es}}
}

@@

The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair. In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian. The former pair, while not producing output suitable for post-editing, is a good example of an assimilation MT system. In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation. For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically. This is not used in this evaluation, but it's still in the toolkit. Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\blankout{\url{https://github.com/Sereni/Appraise}}}

@@

\section{Conclusions}
\label{sec:conclusion}
We have conducted assimilation evaluation of two Apertium translation directions: Basque--Spanish and Tatar--Russian. The results suggest that this evaluation method reflects the contribution of MT to users' understanding of text. The version of the toolkit used in this 
experiment may be downloaded from our repository.\footnote{\blankout{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}

>>>>

\author{{Ekaterina Ageeva}\\
  {School of Linguistics}\\
  {Higher School of Economics}\\
  {Moscow, Russia}\\
  {{\tt evageeva\_2@edu.hse.ru}}\\[2ex]
  {\textbf{Francis M. Tyers}}\\
  {HSL-fakultetet}\\ 
  {UiT Norgga \'{a}rktala\v{s} universitehta} \\
  {9017 Romsa, Norway} \\
  {{\tt francis.tyers@uit.no}}
  \And
  {Mikel L. Forcada}\\
  {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
  \blankout{Universitat d'Alacant, Spain} \\
  {{\tt mlf@dlsi.ua.es}}  \\[2ex]
  {\textbf{Juan Antonio P\'{e}rez-Ortiz}} \\
  {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
  {Universitat d'Alacant, Spain} \\
  {{\tt japerez@dlsi.ua.es}}
}

@@

The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair. In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian. The former pair, while not producing output suitable for post-editing, is a good example of an assimilation MT system. In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation. For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme. In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain. To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically. This is not used in this evaluation, but it's still in the toolkit. Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results. This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{{\url{https://github.com/Sereni/Appraise}}}

@@

\section{Conclusions}
\label{sec:conclusion}
We have conducted assimilation evaluation of two Apertium translation directions: Basque--Spanish and Tatar--Russian. The results suggest that this evaluation method reflects the contribution of MT to users' understanding of text. The version of the toolkit used in this 
experiment may be downloaded from our repository.\footnote{{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
